---
title: "nla2007_census"
author: "Bryan Milstead"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(stringsAsFactors = FALSE) 
library(knitr)
library(tidyverse)
library(RODBC)  #NOTE you must use 32 bit R for this to work

```

## Introduction

* This document ([nla2007_census.pdf](https://github.com/willbmisled/OddAndEnds/blob/master/nla2007_census.pdf)) describes the process of extracting the 2000 census data for the 2007 National Lake Assessment Sites.
* The data were extracted from the 2000 census with the rscript [PopulationLakes_20130212.r](https://github.com/willbmisled/mapping/blob/master/R/PopulationLakes_20130212.r)
* For each lake the number of people and population density were calculated for four buffers.  The buffer sizes were:
    - 300m
    - 1000m
    - 2500m
    - BufRadius: a variable width buffer corresponding to the radius of the lake (see the field BufWidthM).
* The table below shows the data definitions

**Field**  |**units**  | **Definition**
---------------- | ------------- | ------------------------------------------------------------------------------------------ 
NLA_ID | |NLA2007 unique identifier
BufWidthM |m |width of lake buffer for calculations
Pop | people |number of people in the buffer
PopDensityKm2 |People/Km2 | Pop/BufferAreaKm2Adj 
BufferAreaKm2 |Km2 |Area of entire buffer-number of grid cells*30*30/1000000 
BufferAreaKm2Adj |Km2 |Area of Non-NA buffer-number of grid cells with data*30*30/1000000 
PercentNA |% |(number of NA grid cells)/(total number of grid cells)
BufType | |"Fixed" = standard buffer width; "Radius" buffer width = radius of a circle with Area = lake area.

* data were saved in csv format and can be downloaded from github: [nla_2007_census.csv](https://github.com/willbmisled/OddAndEnds/blob/master/nla_2007_census.csv)

```{r getCensusdata, include=FALSE}
# Read data-****Make Sure the Path Is Correct****
con <- odbcConnectAccess("L:/Public/Milstead_Lakes/WaterbodyDatabase/WaterbodyDatabase.mdb")
census <- sqlQuery(con, "
SELECT tblJoinNLAID_WBID.NLA_ID, LakeCensus2000Population.BufWidthM, LakeCensus2000Population.Pop, LakeCensus2000Population.PopDensityKm2, LakeCensus2000Population.BufferAreaKm2, LakeCensus2000Population.BufferAreaKm2Adj, LakeCensus2000Population.PercentNA, LakeCensus2000Population.BufType
FROM tblJoinNLAID_WBID INNER JOIN LakeCensus2000Population ON tblJoinNLAID_WBID.WB_ID = LakeCensus2000Population.WB_ID;
")
close(con)
str(census)

write.csv(census,'nla_2007_census.csv',row.names=FALSE)

```
